The Fates of Merging Supermassive Black Holes and a Proposal for a New Class of X-Ray Sources

### 2020-04-04 paper.tex
FINAL EDITS BETWEEN ME AND JERRY BEFORE SENDING TO ANDREAS FOR HIS FINAL EDITS OR APPROVAL.

\begin{figure}
\vspace{20pt}%
\begin{center}
\includegraphics[width=0.5\textwidth]{plots/"Lx of our Galaxies compared to Babyk data".png}
\caption{X-ray luminosity of hot gas in our galaxies, compared to those from \citet{2018ApJ...857...32B}.}
\label{fig:galgas1}
\end{center}
\end{figure}

@@

\subsection{Galaxy Gas X-Ray Luminosity}
First, we construct realistic X-ray luminosities for our host galaxies. We compute the X-ray luminosity of the hot gas in these galaxies, adapting Equation 26 from \citet{2012ApJ...754..125C}. In lieu of summing over discrete particles, we use the beta model of the gas as defined in our equation \ref{beta_model} and integrate:
\begin{equation}
\begin{aligned}
    L_x ={} \frac{1.2 \times 10^{-24}}{(\mu m_p)^2}\left(\frac{kT_x}{1keV}\right)^{1/2}\rho_0^{2}\;4{\pi} \\
    \int_{0}^{\infty}\frac{r^2}{(1+r^2/r_c^2)^{3}}\;dr\text{  erg s}^{-1}.
\end{aligned}
\end{equation}

@@

\subsection{Black Hole X-Ray Luminosity}
\subsubsection{Assumptions and Method Employed}
Since we model the black holes in this work as point particles for purposes of the simulations, we do not assume any particular structure to them.  Therefore, we adopt the approach and assumptions from \citealt{2018MNRAS.476.1412I} and \citealt{2019MNRAS.486.5377I} in estimating the X-ray luminosities of the remaining black holes left wandering in our subject galaxies.  Their approach assumes rotating accretion flows in a very low-density environment.  Shear viscous forces transport angular momentum outwards, resulting in an accretion disc that forms at the bottom of the distribution.  However, due to the low density of the gas, it does not cool, and convection limits the net accretion, and subsequent radiation luminosity, of the black hole.  In such a convective infall, the gas finally accreted onto the black hole is much less than the initially infalling gravitationally bound gas.

@@

However, due to convective (i.e., inefficient) accretion interior to the Bondi radius (see \cite{2018MNRAS.476.1412I} and \cite{2019MNRAS.486.5377I}), the net accretion to the black hole is often substantially lower than the Bondi-Hoyle accretion.  It can be inferred from the fitting relation given in Equation 31 of \cite{2019MNRAS.486.5377I}:
\begin{subequations}
    \begin{equation}\label{inayoshi_accretion}
        \frac{\dot{M}_{\bullet}}{\dot{M}_{Edd}} \simeq 1.5\times10^{-6}T_7^{-4/5}\times\left(\frac{\alpha}{0.01}\right)^{0.37}\left(\frac{\dot{m}_B}{10^{-3}}\right)^{3/5}\left(\frac{f_c}{0.1}\right)^{2/5},
    \end{equation}
    \begin{equation}\label{mdotb}
        \dot{m}_{B} = \dot{M}_{B}/\dot{M}_{Edd},
    \end{equation}
    \begin{equation}\label{edd_acc}
        \dot{M}_{Edd} = \frac{L_{Edd}}{0.1c^2} \text{  g} \text{ s}^{-1},
    \end{equation}
    \begin{equation}\label{edd_lum}
        L_{Edd} = 1.26\times10^{38}\left(\frac{M_{BH}}{M_{\odot}}\right)\;\text{erg}\;\text{s}^{-1}.
    \end{equation}
\end{subequations}
$\dot{M}_{\bullet}$ is the net accretion to the black hole, $T_7$ is the temperature of the ambient gas in units of $10^7\;K$, $\alpha$ is the strength of viscosity (generally accepted to be 0.01; see \citet{1973A&A....24..337S}, \citet{1991ApJ...376..214B}, \citet{1995ApJ...445..767M}, \citet{1996ApJ...463..656S}, \citet{1998RvMP...70....1B}, \citet{2004PThPS.155..409S}, \cite{2018MNRAS.476.1412I}), and $f_c$ is the conductivity suppression factor (generally accepted to be 0.1; see \citet{2001ApJ...562L.129N}, \citet{2004PhRvL..92d5001M}, and \cite{2019MNRAS.486.5377I}).  The Eddington accretion rate is $\dot{M}_{Edd}$, $c$ is the speed of light, and $L_{Edd}$ is the Eddington luminosity.

@@

Figure \ref{fig:wui} shows the difference in luminosity when taking into account the influence of the central black hole for the 17 wandering black holes that were within the influence radius.  In all cases, the final luminosity is higher than the original luminosity.
\begin{figure}
\begin{center}
\includegraphics[width=0.5\textwidth]{plots/"Comparison of L_x versus radius with and without BH Influence".png}
\caption{Difference in X-ray luminosity of wandering black holes within the influence radius of the central black hole.  For each vertical line, red (lower) dots are the originally calculated luminosity, not taking the CBH influence into account.  Green (upper) dots are the final luminosities, taking the central BH influence into account.}
\label{fig:wui}
\end{center}
\end{figure}

@@

Figure \ref{fig:ldobhs} represents a histogram of the final X-ray luminosities of the black holes, overlaid by a lognormal distribution, obtained with a mean $\mu=1.0\times10^{33}$ and standard deviation $\sigma=4.5\times10^{3}$.
\begin{figure}
\begin{center}
\includegraphics[width=0.5\textwidth]{plots/"Final Luminosities using Kohei factor".png}
\caption{X-ray Luminosity Distribution of Orbiting BHs at z=0.  Green region under curve represents approximate range of detectable orbiting black holes. The dashed, red vertical line is the median X-ray luminosity of our sources ($1.0\times10^{33}\;\text{erg}\;\text{s}^{-1}$).}
\label{fig:ldobhs}
\end{center}
\end{figure}

@@

\section{Conclusions}\label{sec:conclusions}
We performed N-body simulations on 11 galaxies ranging in mass between $10^{12}$ M$_{\sun}$ < M < 3$\times$ $10^{13}$ M$_{\sun}$, from 4 $\geq$ z $\geq$ 0.  A total of 86 orbiting black holes were followed in our simulations, resulting in 42$\%$ mergers with their host galaxy's resident black hole, while another 43$\%$ remained in orbit at $z=0$.  Those remaining orbiting black holes had X-ray luminosities in the range of $10^{28}-10^{43}\mathrm{erg}~\mathrm{s}^{-1}$.  The median luminosity of $10^{33}\mathrm{erg}~\mathrm{s}^{-1}$ would be undetectable with present instruments, although the top $5\%$ most luminous from this set could be detectable.

>>>>

\begin{figure}
\vspace{20pt}%
\begin{center}
\includegraphics[width=0.5\textwidth]{plots/"Lx of our Galaxies compared to Babyk data".png}
\caption{X-ray thermal luminosity of hot gas in our galaxies, compared to those from \citet{2018ApJ...857...32B}.}
\label{fig:galgas1}
\end{center}
\end{figure}

@@

\subsection{Galaxy Gas X-Ray Thermal Luminosity}
First, we construct realistic X-ray thermal luminosities for our host galaxies. We compute the X-ray thermal luminosity of the hot gas in these galaxies, adapting Equation 26 from \citet{2012ApJ...754..125C}. In lieu of summing over discrete particles, we use the beta model of the gas as defined in our equation \ref{beta_model} and integrate:
\begin{equation}
\begin{aligned}
    L_x ={} \frac{1.2 \times 10^{-24}}{(\mu m_p)^2}\left(\frac{kT_x}{1keV}\right)^{1/2}\rho_0^{2}\;4{\pi} \\
    \int_{0}^{\infty}\frac{r^2}{(1+r^2/r_c^2)^{3}}\;dr\text{  erg s}^{-1}.
\end{aligned}
\end{equation}

@@

\subsection{Black Hole X-Ray Thermal Luminosity}
\subsubsection{Assumptions and Method Employed}
Since we model the black holes in this work as point particles for purposes of the simulations, we do not assume any particular structure to them.  Therefore, we adopt the approach and assumptions from \citealt{2018MNRAS.476.1412I} and \citealt{2019MNRAS.486.5377I} in estimating the X-ray luminosities of the remaining black holes left wandering in our subject galaxies.  Their approach assumes rotating accretion flows in a very low-density environment.  Shear viscous forces transport angular momentum outwards, resulting in an accretion disc that forms at the bottom of the distribution.  However, due to the low density of the gas, it does not cool, and convection limits the net accretion, and subsequent radiation luminosity, of the black hole.  In such a convective infall, the gas finally accreted onto the black hole is much less than the initially infalling gravitationally bound gas.

@@

However, due to convective (i.e., inefficient) accretion interior to the Bondi radius (see \cite{2018MNRAS.476.1412I} and \cite{2019MNRAS.486.5377I}), the net accretion to the black hole is often substantially lower than the Bondi-Hoyle accretion.  It can be inferred from the fitting relation given in Equation 31 of \cite{2019MNRAS.486.5377I}:
\begin{subequations}
    \begin{equation}\label{inayoshi_accretion}
        \frac{\dot{M}_{\bullet}}{\dot{M}_{Edd}} \simeq 1.5\times10^{-6}T_7^{-4/5}\times\left(\frac{\alpha}{0.01}\right)^{0.37}\left(\frac{\dot{m}_B}{10^{-3}}\right)^{3/5}\left(\frac{f_c}{0.1}\right)^{2/5},
    \end{equation}
    \begin{equation}\label{mdotb}
        \dot{m}_{B} \equiv \dot{M}_{B}/\dot{M}_{Edd},
    \end{equation}
    \begin{equation}\label{edd_acc}
        \dot{M}_{Edd} \equiv \frac{L_{Edd}}{0.1c^2} \text{  g} \text{ s}^{-1},
    \end{equation}
    \begin{equation}\label{edd_lum}
        L_{Edd} \equiv 1.26\times10^{38}\left(\frac{M_{BH}}{M_{\odot}}\right)\;\text{erg}\;\text{s}^{-1}.
    \end{equation}
\end{subequations}
$\dot{M}_{\bullet}$ is the net accretion to the black hole, $T_7$ is the temperature of the ambient gas in units of $10^7\;K$, $\alpha$ is the strength of viscosity (generally accepted to be 0.01; see \citet{1973A&A....24..337S}, \citet{1991ApJ...376..214B}, \citet{1995ApJ...445..767M}, \citet{1996ApJ...463..656S}, \citet{1998RvMP...70....1B}, \citet{2004PThPS.155..409S}, \cite{2018MNRAS.476.1412I}), and $f_c$ is the conductivity suppression factor (generally accepted to be 0.1; see \citet{2001ApJ...562L.129N}, \citet{2004PhRvL..92d5001M}, and \cite{2019MNRAS.486.5377I}).  The Eddington accretion rate is $\dot{M}_{Edd}$, $c$ is the speed of light, and $L_{Edd}$ is the Eddington luminosity.

@@

Figure \ref{fig:wui} shows the difference in luminosity when taking into account the influence of the central black hole for the 17 wandering black holes that were within the influence radius.  In all cases, the final luminosity is higher than the original luminosity.
\begin{figure}
\begin{center}
\includegraphics[width=0.5\textwidth]{plots/"Comparison of L_x versus radius with and without BH Influence".png}
\caption{Difference in X-ray thermal luminosity of wandering black holes within the influence radius of the central black hole.  For each vertical line, red (lower) dots are the originally calculated luminosity, not taking the CBH influence into account.  Green (upper) dots are the final luminosities, taking the central BH influence into account.}
\label{fig:wui}
\end{center}
\end{figure}

@@

Figure \ref{fig:ldobhs} represents a histogram of the final X-ray luminosities of the black holes, overlaid by a lognormal distribution, obtained with a mean $\mu=1.0\times10^{33}$ and standard deviation $\sigma=4.5\times10^{3}$.
\begin{figure}
\begin{center}
\includegraphics[width=0.5\textwidth]{plots/"Final Luminosities using Kohei factor".png}
\caption{X-ray Thermal Luminosity Distribution of Orbiting BHs at z=0.  Green region under curve represents approximate range of detectable orbiting black holes. The dashed, red vertical line is the median X-ray thermal luminosity of our sources ($1.0\times10^{33}\;\text{erg}\;\text{s}^{-1}$).}
\label{fig:ldobhs}
\end{center}
\end{figure}

@@

\section{Conclusions}\label{sec:conclusions}
We performed N-body simulations on 11 galaxies ranging in mass between $10^{12}$ M$_{\sun}$ < M < 3$\times$ $10^{13}$ M$_{\sun}$, from 4 $\geq$ z $\geq$ 0.  A total of 86 orbiting black holes were followed in our simulations, resulting in 42$\%$ mergers with their host galaxy's resident black hole, while another 43$\%$ remained in orbit at $z=0$.  Those remaining orbiting black holes had X-ray luminosities in the range of $10^{28}-10^{43}\mathrm{erg}~\mathrm{s}^{-1}$.  The median luminosity of $10^{33}\mathrm{erg}~\mathrm{s}^{-1}$ would be undetectable with present instruments, although the top $5\%$ most luminous from this set could be detectable.  If detected, these orbiting massive black holes would constitute a new category of X-ray source.
